Comparing Reinforcement Learning and Finite State Machine Agents in Real Time Strategy Games: Impact on Player Experience

### 2025-03-31 Paper/main.tex
added todos, notes and parts to the paper in preparation for when Sem 1 assignments are almost over, and enough time can be dedicated to Sem 2 assignments.

@@

@@

>>>>

% The difference between academic and non academic literature is that academic literature is peer-reviewed, and is as such, more reliable and trustworthy
% than non-academic literature, which can easily be biased or contain false information. Academic literature can also be more in-depth and detailed, due to
% the high research standards and requirements of academic institutions, especially IEEE.

% TODO: Filter through the papers in zotero and remove the excessive/unnecessary ones while developing the literature map and review.

@@

% Mention Csikszentmihalyi's flow theory and how it relates to player experience and immersion in games.

The goal of a game developer is to create a game that is fun and engaging, and they do this by creating a game that is challenging, but not too challenging.
As Csikszentmihalyi's flow theory states, the goal of a game is to create a state of flow, where the player is fully immersed in the game, and is not
bored, but also not overwhelmed. There are many ways to achieve this, and one of them is through the use of AI opponents. Traditionally, FSMs are
used as AI opponents, as they are simple to implement and easy to understand, and when done correctly, provide a good challenge to the player.
However, given enough time, players can learn the patterns of the FSMs, and exploit them, making the game feel boring. It is possible to combat this
through weakening the player, or making the AI more difficult, as is done in Souls-like games, however this can prove too challenging and overwhelm the player,
once again breaking the flow.

RL agents, on the other hand, are able to learn how to play the game, and as such, adapt to the player and given scenario. This creates a more engaging
and immersive experience, as the player feels like they are playing against a real opponent, and not just a computer. This is especially true in RTS games,
due to the complexity of the game, and the many strategies that can be taken, which can be seen in the work done by \cite{vinyals_grandmaster_2019}.
There is an issue with this, however, as not only are RL Agents computationally expensive, but they can be trained too far, and become too difficult for the player,
once again breaking the flow state. In the work done by \cite{grech_creating_2023}, they balance this by taking snapshots of the RL agent at different
stages in training. This allows them to create different difficulties of AI opponents, which can be used to better match a player's skill level.

% Another way of doing this, is by building dynamic difficulty adjustment (DDA) systems into the opponent, which work by adjusting the rewards and punishments
% of the agent to make them easier or harder to play against (Add citation from zotero.).

% Go about comparing these 3 different approaches, and how they always compare the performance of the RL agent, and not the player experience.

@@

% Research Questions:

% 1. How do RL and FSM AI opponents compare in terms of player experience in RTS games?
% 2. What are the key factors that influence player experience when playing against RL and FSM AI opponents in RTS games?
% 3. How do RL and FSM AI opponents impact player immersion in RTS games?

% Research objective:

% The objective for the research is to evaluate the impact of RL and FSM opponents in RTS games, and to determin their impact on player experience and immersion,
% and what are the key factors that cause this impact/infuence. This is to be done by creating a simple RTS game, and implementing both RL and FSM agents,
% and then conducting a play test experiment with players, where both groups will then be surveyed to gather data on their experience. Along with this,
% data gathered during the playtest through unity analytics will be used to measure player engagement and immersion.
